---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(readr)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

```{r}
str(av)
summary(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers.

```{r}
deaths <- av %>%
  select(Name.Alias, starts_with("Death")) %>%
  pivot_longer(
    cols = starts_with("Death"),
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(
    Time = parse_number(Time),
    Death = tolower(trimws(Death))
  )         

head(deaths)
tail(deaths)
str(deaths)
deaths
```
```{r}
returns <- av %>%
  pivot_longer(cols = starts_with("Return"),
               names_to = "Time",
               values_to = "Return") %>%
  mutate(Time = parse_number(Time),
         Return = tolower(Return))

head(returns)
tail(returns)
str(returns)
```

```{r}
avg_deaths <- deaths %>%
  group_by(Name.Alias) %>%
  summarize(num_deaths = sum(Death =="yes", na.rm = TRUE)) %>%
  ungroup() %>%
  summarize(average_deaths = mean(num_deaths))

avg_deaths
```

```{r}
debug_deaths <- deaths %>%
  group_by(Name.Alias) %>%
  summarize(num_deaths = sum(Death == "yes", na.rm = TRUE))

print(debug_deaths)
```
## On average each Avenger experiences 0.5460123 deaths.

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

#Lucas -> "There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death."

### Include the code

#Lucas ->
```{r}
av_long <- av %>%
   pivot_longer(cols = matches("Death[1-5]|Return[1-5]"),
                names_to = c(".value", "Time"),
                names_pattern = "(Death|Return)([1-5])") %>%
  mutate(across(c(Death, Return), tolower),
         Time = as.numeric(Time))

first_return_rate <- av_long %>%
  filter(Time == 1, Death == "yes") %>%
  summarize(
    total_died_first = n(),
    total_returned_first = sum(Return == "yes", na.rm = TRUE),
    percent_returned_first = (total_returned_first  / total_died_first) * 100
    )

second_third_return_rate <- av_long %>%
  filter(Time %in% c(2,3), Death == "yes") %>%
  summarize(
    total_died_second_third = n(),
    total_returned_second_third = sum(Return == "yes", na.rm = TRUE),
    percent_returned_second_third = (total_returned_second_third / total_died_second_third) * 100
  )

first_return_rate
second_third_return_rate
```

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.
